\scuba: iterative map-making with the Sub-Millimetre User Reduction Facility

### 2012-11-22 smurf_paper.tex
more explanation about why zeromasking works. fixed m17 figs

\begin{figure}
\centering
\includegraphics[width=\linewidth]{com}
\caption{A selection of typical bolometer time-series after
flatfielding and removal of a constant baseline. Each sub-plot shows
the power from a single bolometer over a 30 second period.
The lower and upper vertical limits of each plot are -0.02 pW and +0.02
pW respectively. It can be seen that most bolometers exhibit a common
time variation overlaid with other features. Empty squares indicate the
locations of broken bolometers.}
\label{fig:com}
\end{figure}

@@

The bolometer residual variances, $\sigma^2_i$, are measured once the
modelled signal components have been removed. In the early iterations,
the residuals contain both white noise, and other long-timescale
features. However, when the solution has converged, this signal should
look white. Taking the component of bolometer variances due to white
noise, $\sigma_i^{\mathbf{w}2}$ (measured from the bolometer PSDs in
\model{NOI}, Section~\ref{sec:noi}), $\chi^2_\mathrm{r}$ is calculated
as $(1/N) \sum_i \sigma^2_i / \sigma_i^{\mathbf{w}2}$, where $i$ runs
over the $N$ bolometers. In other words, it is the average ratio
between the measured time-series bolometer variances and their
white-noise levels measured between 2 and 10\,Hz, and should tend to a
value of $\sim$1 if the low-frequency noise (and any bright
astronomical signals) have been successfully removed. However, this
expression does not account for the degrees of freedom in the
model. Clearly there are a large number of parameters that will
``fit-out'' some of the uncorrelated white noise, and bias this
estimate of $\chi^2_\mathrm{r}$ low. Indeed, for a converged map,
values near $\sim$0.85 are common, depending primarily on the fraction
of the total bandwidth that has been removed by high-pass filtering
(Section~\ref{sec:flt}). Regardless of the normalisation, once the
model parameterisation has been chosen, this quantity should converge
to some fixed value.

@@

While most of the steps described in the Section~\ref{sec:algorithm}
for reducing data with SMURF are applicable to all data sets, there is
no ``default'' reduction that provides good results for a wide range
of source \snr\ and angular scales. However, SMURF does have a small
number of configurations with variations on these baseline parameters
that are applicable to most common types of data. In this section we
illustrate the differences bewteen these configurations with the
following examples: a bright point source, Uranus
(Section~\ref{sec:point}); a blind survey of high-redshift galaxies in
the Lockman Hole (Section~\ref{sec:cosmo}); and a map of a bright
extended star-forming region in our Galaxy, M17
(Section~\ref{sec:extended}).

@@

This example illustrates the need for constraints in the map solution
in many situations. For calibrators (and other previously known
bright, compact sources), a good, simple prior is to constrain the map
to a value of zero away from the known locations of emission (part of
the \model{AST} model calculation, Section~\ref{sec:ast}). In
Fig.~\ref{fig:pointmaps}d, a solution is produced in an identical
manner to Fig.~\ref{fig:pointmaps}c, but now setting the map
explicitly to zero beyond a radius of 60\,arcsec from the location of
Uranus (much larger than the FWHM of the main lobe), for all but the
final iteration. In this case, the map converges after 6 iterations,
and the ringing has been effectively removed. Since the map is now
flat away from the source, and constrained to a value of zero, it is
appropriate for performing aperture photometry directly, with no need
for an additional reference annulus. This approach to constrained
map-making is similar to that employed for poorly cross-linked scans
of compact (though resolved) sources by \citet{wiebe2009} using BLAST
data.

@@

\includegraphics[width=\linewidth]{cor_pspec_m17_default}
\end{minipage}
\caption{Angular PSDs for the region of the M17 map in
Fig.~\ref{fig:m17_jk} containing synthetic data, using the baseline
configuration. (a) Raw PSDs for the input (noiseless) simulated data
(thick black line), the signal (average of each half) PSDs (thin solid
lines), and noise PSDs estimated from the jackknives (dashed
lines). Vertical dotted lines indicate the high-pass filter scales:
150\,arcsec (red); 300\,arcsec (orange); 600\,arcsec (green); and
900\,arcsec (blue). PSDs are also colour-coded by filter scale. (b)
Since the input PSD is known, it is possible to measure the transfer
function of the map-maker as the ratio between the difference of the
output map signal PSDs and jackknife PSDs, and the input PSDs, giving
the thin coloured lines (linear vertical axis shown on right of
plot). The remaining lines are as in (a), but now corrected by the
transfer function. This plot shows that increasing the filter scale
improves the \snr\ at intermediates scales ($\sim$200--600\,arcsec),
although the \snr\ worsens at smaller scales ($\lsim$200\,arcsec).}
\label{fig:m17_def_ps}
\end{figure*}

>>>>

\begin{figure}
\centering
\includegraphics[width=\linewidth]{com}
\caption{A selection of bolometer time-series from a representative
region of the focal plane after flatfielding and removal of a constant
baseline. Each sub-plot shows the power from a single bolometer over a
30\,s interval.  The lower and upper vertical limits of each plot are
$-0.02$\,pW and $+0.02$\,pW, respectively. It can be seen that most
bolometers exhibit a common time variation overlaid with other
features. Empty squares indicate the locations of broken bolometers.}
\label{fig:com}
\end{figure}

@@

The bolometer residual variances, $\sigma^2_i$, are measured once the
modelled signal components have been removed. In the early iterations,
the residuals contain both white noise, and other long-timescale
features. However, when the solution has converged, this signal should
look white. Taking the component of bolometer variances due to white
noise, $\sigma_i^{\mathbf{w}2}$ (measured from the bolometer PSDs in
\model{NOI}, Section~\ref{sec:noi}), $\chi^2_\mathrm{r}$ is calculated
as $(1/N) \sum_i \sigma^2_i / \sigma_i^{\mathbf{w}2}$, where $i$ runs
over the $N$ bolometers. In other words, it is the average ratio
between the measured time-series bolometer variances and their
white-noise levels measured between 2 and 10\,Hz, and should tend to a
value of $\sim$1 if the low-frequency noise (and any bright
astronomical signals) have been successfully removed. However, this
expression does not account for the degrees of freedom in the
model. Clearly there are a large number of parameters that will
``fit-out'' some of the uncorrelated white noise, and bias this
estimate of $\chi^2_\mathrm{r}$ low. Indeed, for a converged map,
values in the range $\sim$0.85--0.95 are common, depending on the
precise configuration parameters. Regardless of the normalisation,
once the model parameterisation has been chosen, this quantity should
converge to some fixed value.

@@

While most of the steps described in Section~\ref{sec:algorithm} for
reducing data with SMURF are applicable to all data sets, there is no
``default'' reduction that provides good results for a wide range of
source \snr\ and angular scales. However, SMURF does have a small
number of configurations with variations on these baseline parameters
that are applicable to most common types of data. In this section we
illustrate the differences bewteen these configurations with the
following examples: a bright point source, Uranus
(Section~\ref{sec:point}); a blind survey of high-redshift galaxies in
the Lockman Hole (Section~\ref{sec:cosmo}); and a map of a bright
extended star-forming region in our Galaxy, M17
(Section~\ref{sec:extended}).

@@

This example illustrates the need for constraints in the map solution
in many situations. For calibrators (and other previously known
bright, compact sources), a good, simple prior is to constrain the map
to a value of zero away from the known locations of emission (part of
the \model{AST} model calculation, Section~\ref{sec:ast}).  In
Fig.~\ref{fig:pointmaps}d, a solution is produced in an identical
manner to Fig.~\ref{fig:pointmaps}c, but now setting the map
explicitly to zero beyond a radius of 60\,arcsec from the location of
Uranus (much larger than the FWHM of the main lobe), for all but the
final iteration. In this case, the map converges after 6 iterations,
and the ringing has been effectively removed. Since the map is now
flat away from the source, and constrained to a value of zero, it is
appropriate for performing aperture photometry directly, with no need
for an additional reference annulus.

The reason this prior works is simple to understand if the bolometer
data are thought of as providing useful differential measurements up
to scales corresponding to the filter edge. In this example, the
bolometers retain information on brightness differences up to
200\,arcsec scales. Since the map is constrained to zero to within
60\,arcsec of the peak of Uranus (well within 200\,arcsec), the
solution is then able to accurately reconstruct the differential peak
intensity of Uranus with respect to this constrained background.  This
approach to map-making is similar to that employed for poorly
cross-linked scans of compact (though resolved) sources by
\citet{wiebe2009} using BLAST data.

@@

\includegraphics[width=\linewidth]{cor_pspec_m17_default}
\end{minipage}
\caption{Angular PSDs for the region of the M17 map in
Fig.~\ref{fig:m17_jk} containing synthetic data, using the baseline
configuration. (a) Raw PSDs for the input (noiseless) simulated data
(thick black line), the signal (average of each half) PSDs (thin solid
lines), and noise PSDs estimated from the jackknives (dashed
lines). Vertical dotted lines indicate the high-pass filter scales:
150\,arcsec (red); 300\,arcsec (orange); 600\,arcsec (green); and
900\,arcsec (blue). PSDs are also colour-coded by filter scale. (b)
Since the input PSD is known, it is possible to measure the transfer
function of the map-maker as the ratio between the difference of the
output map signal PSDs and jackknife PSDs, and the input PSDs, giving
the thin coloured lines (linear vertical axis shown on right of
plot). The remaining lines are as in (a), but now corrected by the
transfer function. This plot shows that increasing the filter scale
improves the \snr\ at intermediate scales ($\sim$200--600\,arcsec),
although the \snr\ worsens at smaller scales ($\lsim$200\,arcsec).}
\label{fig:m17_def_ps}
\end{figure*}
